---
title: "salmon_group25"
author: "Laurel Wee, Cali Pfleger, Carmen Galaz-García"
date: "1/29/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(tsibble) # time series tibbles
library(feasts)
library(patchwork)
```

### Overview

- picture
- Summary
- map

## Data Citation

```{r, basic data set, warning=FALSE}
fish <- read_csv(here("data","willamette_fish_passage.csv")) %>% 
  clean_names() %>% 
  select(date,steelhead,coho,jack_coho) %>% 
  mutate(date = lubridate::mdy(date)) %>%     # convert to date type
  rename(jack = jack_coho) %>%                # update column name
  pivot_longer(cols = steelhead:jack,         # convert to tidy data
               names_to="fish_type", 
               values_to = "n_crossing") %>% 
  as_tsibble(key=fish_type, index= date)           # convert date into tsibble
```


## Willamette Falls fish passage {.tabset}


### Original time series
```{r, tab 1}

```

### Seasonplots
```{r, tab 2, out.width="100%"}

basic_seasonplot <- function(fish_data, fish_type_str){
 
  month_fish_plot <- fish_data %>% 
    filter(fish_type == fish_type_str) %>%  
    index_by(yr_mo = ~yearweek(.)) %>% # group by year and month
    summarize(monthly_mean_crossings = mean(n_crossing, na.rm=TRUE))  %>% 
    gg_season(y=monthly_mean_crossings) +
    labs(x=" ",
         y="weekly average passage")+
    theme_light()
  
  return(month_fish_plot)
}

steel_season <- basic_seasonplot(fish, "steelhead")+
  labs(title = "Graph 1. Adult Steelhead  Passage")

coho_season <- basic_seasonplot(fish, "coho") +
  labs(title = "Graph 2. \nAdult Coho Passage")

jack_season <- basic_seasonplot(fish, "jack") +
  labs(title = "Graph 3. \nJack Coho Passage")+
  labs(y="")




steel_season / (coho_season| jack_season) +
  plot_layout(guides = "collect") & theme(legend.position = 'right')
```

**Figure 2.** 
Average weekly count of three groups of salmon passing through the Willamette Falls fish ladder from 2001 to 2010. 
Graph 1 shows the average weekly passage by year of adult steelhead salmon (*oncorhynchus mykiss*).
Graphs 2 and 3 show, respectively, the average weekly passage by year of adult and jack coho salmon (*oncorhynchus kisutch*).  
Data: Columbia River Dart.

**Trends:**

- For adult steelhead salmon we can observe clear seasonaility of the weekly average passage. 
The  average count increases from December to May/June, then decreases until August/September and remains in less than 100 average crossings until December.

- Both groups of coho salmon have the same seasonality in weekly average passage. Coho salmon seem to only use the fish ladder from August to December regardless of life stage, with crossings increasing from August and peaking in September. 

- Average weekly crossings of adult coho salmon have increased from 2001 to 2010, while passage of adult steelhead salmon has decreased in the same period. 

### Annual counts by species
```{r, tab 3}

```


